---
title: "HW1"
author: "Team 1"
date: "September 13, 2020"
output:
  pdf_document: default
  html_document: default
---

```{r}
# load required packages
library(ggplot2)
library(dplyr)
#library(tidyr)
library(MASS)
library(caret)
library(RCurl)
```

```{r import}
# Loading the data
git_dir <- 'https://raw.github.com/odonnell31/data621-HW1/master/data'
train_df = read.csv(paste(git_dir, "/moneyball-training-data.csv", sep=""))
test_df = read.csv(paste(git_dir, "/moneyball-evaluation-data.csv", sep = ""))
```

# 1. Data Exploration

See a summary of each column in the train_dfing set
```{r train_dfing_data_summary}
# view a summary of all columns
summary(train_df)
```

For types of hits, see a histogram of each
```{r hits_histograms}
par(mfrow=c(2,2))
hist(train_df$TEAM_BATTING_H,
     main = "hits histogram", xlab = "hits (season)",
     breaks = 20)
hist(train_df$TEAM_BATTING_2B,
     main = "doubles histogram", xlab = "doubles (season)",
     breaks = 20)
hist(train_df$TEAM_BATTING_3B,
     main = "triples histogram", xlab = "triples (season)",
     breaks = 20)
hist(train_df$TEAM_BATTING_HR,
     main = "homeruns histogram", xlab = "homeruns (season)",
     breaks = 20)
par(mfrow=c(1,1))
```

```{r}
pairs(~ TARGET_WINS + TEAM_BATTING_H + TEAM_BATTING_2B
      + TEAM_BATTING_3B + TEAM_BATTING_HR, data = train_df)
```
```{r}
# look at the structure of the variables
str(train_df)
str(eval)
```

```{r}
# lets observe how targets_win are effected by other factors
hist(train_df$TARGET_WINS,xlab="TARGET_WINS",main="")
# we have no TARGET_WINS from eval
# hist(eval$TARGET_WINS,xlab="TARGET_WINS",main="")
```



# 2. Data Preparation

1. We are told everything is standardized to match a 162 game season, so it is my preference to make TARGET_WINS a decimal of 162
```{r}
train_target_wins = train_df$TARGET_WINS
train_df$TARGET_WINS = train_df$TARGET_WINS/162.
# TARGET_WINS now a decimal of games won in 162 game season
hist(train_df$TARGET_WINS,xlab="TARGET_WINS",main="")
str(train_df)
```

2. Assuming that everything that is NA can be filled by 0 based on the description of variables, create columns flagging if original values were NA (e.g. create TEAM_BATTING_HBP_NA column and value is 1 if TEAM_BATTING_HBP is NA and 0 otherwise meaning it wasn't NA and had a value. Do this for all columns)
```{r}
# 
has_NA = names(which(sapply(train_df, anyNA)))
for (col in has_NA)
{
   new_col = (paste(col,"_NA", sep=""))
   train_df[,new_col] = as.numeric(is.na(train_df[,col]))
   test_df[,new_col] = as.numeric(is.na(test_df[,col]))
}
train_df[is.na(train_df)] = 0
test_df[is.na(test_df)] = 0
```

# 3. Build Models
```{r}
# set seed for reproducibility
n_records = nrow(train_df)
set.seed(1)
# Random sample indexes
# train_df_idx <- sample(1:nrow(train_df), 0.9 * nrow(train_df))
# val_idx <- setdiff(1:nrow(train_df), train_df_idx)
# val = train_df[val_idx,]
# train_df = train_df[train_df_idx,]
```

```{r}
#model = lm(TARGET_WINS ~ ., data=train_df)
#summary(model)
# drop TEAM_PITCHING_SO_NA because model thinks its correlated with another var
#train_df = subset(train_df, select = -c(TEAM_PITCHING_SO_NA))
#model = lm(TARGET_WINS ~ ., data=train_df)
#summary(model)
```

```{r}
#layout(matrix(c(1,2,3,4),2,2)) # optional 4 graphs/page
#plot(model)
```

# following ideas for model selecting taken from http://www.sthda.com/english/articles/37-model-selection-essentials-in-r/154-stepwise-regression-essentials-in-r/
```{r}
# Try different selection methods
full_model = lm(TARGET_WINS ~ ., data=train_df)
step.model <- stepAIC(full_model, direction = "both", 
                      trace = FALSE)
summary(step.model)
```

```{r}
# Train model
train_control = trainControl(method = "cv", number = 10)
step_model = train(TARGET_WINS ~ ., data=train_df,
                   method = "lmStepAIC",
                   trControl = train_control,
                   trace=FALSE)
# Model accuracy
step_model$results

# Final model coefficients
step_model$finalModel

# Summary of model
summary(step_model$finalModel)
```

```{r}
model = step_model$finalModel
plot(fitted(model), resid(model))
hist(model$residuals)
qqnorm(resid(model))
qqline(resid(model))
```

# Predictions on Evaluation Set

```{r}
# convert decimals of wins back to number of wins, rounded
test_preds = round(predict(model, newdata=test_df)*162)
test_df$PRED_TARGET_WINS = test_preds
# write out evaluation data with predictions
write.csv(test_df, 'data/eval_with_preds.csv')
hist(test_preds)
hist(train_target_wins)
```